---
title: "Quick Usage Guide"
author: "Colman Humphrey"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick Usage Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_knit$set(cache.extra = 234) # seed
```

For even quicker usage instructions, see the readme.

The general idea of this package is that you can throw nearly any 
model built in the typical R fashion. For now, we just need 
a somewhat standard way of extracting fixed effects, and that
`update` works, which is nearly always the case. 

If your model is called `base_model` for example, then 
if `coef(summary(base_model))` or `coef(summary(base_model))$cond`
gives the estimates and standard errors, the method will work.

The data should be supplied manually, although it's currently the case 
that if the dataframe that creates the model is accessible from
`base_model$model`, `base_model$frame` or `base_model@frame`, you 
won't have to send the data in to the function separately. Be 
careful here - some models supply this but the method introduces bugs. 
Much safer to supply the data explicitly.

Let's work through an example with `glmmTMB`. We'll use the data
that comes with `glmmboot`.

```{r, cache=TRUE}
library(glmmboot)
data(test_data)

head(test_data)
```

We're assuming the `x_var` variables are fixed, and `subj` is to be treated as a random effect.

Thus our base analysis is:

```{r, cache=TRUE}
library(glmmTMB)
model_formula <- as.formula(y ~ x_var1 + x_var2 + x_var2 + (1 | subj))

base_run <- glmmTMB(formula = model_formula,
                    data = test_data,
                    family = binomial)
```
We get a warning because the outcome data is proportional. Not to worry.

Now we'll use the bootstrap. By default it'll perform block bootstrapping over the highest
entropy random effect - but there's only one, so of course the winner is `subj`.

```{r, cache=TRUE}
bootstrap_over_subj <- BootGlmm(base_model = base_run,
                                resamples = 99, 
                                base_data = test_data,
                                num_cores = 4)
```
For publications etc, better to run about ten thousand resamples to avoid
noise having much of an effect. Of course 99 is far too small, only for an example.

And comparing results:
```{r, cache=TRUE}
print(bootstrap_over_subj)
```

This of course might take a long time. If it takes far too long on your machine,
you can ideally run it on a bunch of computers. We don't want each computer to 
output the fully processed output, only the intermediate outcome. To do this, 
we set `return_coefs_instead = TRUE` for each run:
```{r, cache=TRUE}
b_list1 <- BootGlmm(base_model = base_run,
                    resamples = 29, 
                    base_data = test_data,
                    num_cores = 1,
                    return_coefs_instead = TRUE)
b_list2 <- BootGlmm(base_model = base_run,
                    resamples = 30, 
                    base_data = test_data,
                    num_cores = 1,
                    return_coefs_instead = TRUE)
b_list3 <- BootGlmm(base_model = base_run,
                    resamples = 30, 
                    base_data = test_data,
                    num_cores = 1,
                    return_coefs_instead = TRUE)
```
Combining this is simple enough. If we've used a few, we don't want to mess around with 
even more lists, so we can enter them into the relevant function:
```{r, cache=TRUE}
CombineResampledLists(b_list1, b_list2, b_list3)
```

If we've run a huge number of such runs, ideally we'll combine all output to a list of lists, like
so:
```{r, cache=TRUE}
list_of_lists_output <- list(b_list1, b_list2, b_list3)
```
And we'll get the same result:
```{r, cache=TRUE}
CombineResampledLists(list_of_lists_output)
```
<!---
Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
--->
